---
title: "Activity 8"
author: "Robert Bilyk"
output: github_document
---

## Data and packages

Load the entirety of the `{tidyverse}`.
`{forcats}` and `{stringr}` are loaded as part of this.
If you wish to work with dates during this activity, you will need to also load `{lubridate}`.
Be sure to avoid printing out any unnecessary information and give the code chunk a meaningful name.

```{r}
library(tidyverse)
```



Cheatsheets that you might want to add to your collection:

- [`{stringr}`](https://stringr.tidyverse.org/)
- [`{forcats}`](https://forcats.tidyverse.org/)
- [`{lubridate}`](https://lubridate.tidyverse.org/)

Using `here::here`, upload the `billboard_songs.txt` file that is saved in your `data` folder.
Notice that this file is a tab-delimited file that is stored as a `.txt` file.
Therefore, you will need to use either `read_delim` with a `delim = ...` argument or (better) `read_tsv` .
Assign the file to a meaningful object name, be sure to avoid printing any unnecessary information, and give the code chunk a meaningful name.

```{r}
songs <- read_tsv(here::here("data","billboard_songs.txt"))
```



These data include information on song popularity.
In the US, the Billboard Hot 100 is a list that comes out every week, showing the 100 most played songs that week.
More information about the creation of this dataset, as well as some analyses by the author, can be found here: [https://mikekling.com/analyzing-the-billboard-hot-100/](https://mikekling.com/analyzing-the-billboard-hot-100/).
The dataset you are provided is a limited version of the full data, containing:

- `title`
- `artist`
- `overall peak`: The highest rank the song ever reached (1 is the best)
- `weeks on chart`: The number of weeks the song was on the chart
- `chart date`: The latest date the song appeared on the Billboard Hot 100

This is a long dataset (34,605 observations)!
You might like to create a small dataset with only, say, 200 of the rows to try all your code out on the smaller dataset first, and then only run the analysis of the full data after you have experienced everything.
One way to do this is to use a function like `slice_sample(n = ...)` from `{dplyr}`.

```
# Randomly sample 200 rows
```{r}
songs %>% 
  slice_sample(n = 200)
```

If you wish to work with `{stringr}`, I find it useful to work on vectors first.
A way to do this is with `pull` from `{dplyr}`.

```
# Turn variable from dataset into a vector
dataset %>% 
  pull(variable)
```
```{r}
songs %>%
  pull(variable)
```

## Analysis

You are encouraged to explore these data as you wish using functions from the packages for the three special variable types.
Some ideas that you might be interested in:

- What 10 songs (display title, artist, and week) were on the charts for the longest?
```{r}
songs %>%
  arrange(desc(`weeks on chart`))
```

- What distinct date did the oldest song(s) leave the charts?
```{r}
library(lubridate)

songs %>%
  arrange(`chart date`)%>%
  mutate(`chart date` = ymd(`chart date`), `off chart` = `chart date` + weeks(`weeks on chart`))%>%
  print(min(songs$`off chart`))
```

- What songs could have been played at your 16th birthday party?
```{r}
my_birthday <- mdy("11-14-2000")
songs %>%
  arrange(`chart date`)%>%
  mutate(`chart date` = ymd(`chart date`))
```

That is, which songs that eventually peaked at #1  **entered** the charts within a couple months (before or after) your 16th birthday?

Assuming you were 16 years old by 2015.
- Which artist has been **featured** on the most Billboard charting songs?
```{r}
artist <- songs %>%
            pull(artist)
str_art <- artist %>%
  str_split(pattern="featuring") %>%
  unlist()
  
str_table <- tibble(artist=str_art)%>%
  group_by(artist) %>%
  summarise(n=n())%>%
  arrange(desc(n))
print(str_table)
```

- Which artist has **collaborated** on the most Billboard charting songs?
```{r}
collaborate <- str_subset(artist,"featuring")
print(collaborate)
```

- Create some data visualization controlling the order of the character/string variables.

## Attribution

Parts of this Activity are based on a lab from [Dr. Kelly Bodwin's](https://www.kelly-bodwin.com/) STAT 331 course.